% !TEX program = xelatex
\documentclass[12pt,letterpaper]{article}

\usepackage[top=1in, bottom=1in, left=1.1in, right=1.1in]{geometry}

\usepackage{fontspec}
\setmainfont[
  Extension = .otf,
  UprightFont = lmsans10-regular,
  BoldFont = lmsans10-bold,
  ItalicFont = lmsans10-oblique,
  BoldItalicFont = lmsans10-boldoblique,
]{lmsans10}

\newfontfamily\acbold{ywft-processing-bold}[
  Path=../../system/public/type/webfonts/,
  Extension=.ttf
]

\usepackage{xcolor}
\usepackage{tabularx}
\usepackage{fancyhdr}
\usepackage{ragged2e}
\usepackage{microtype}
\usepackage{hyperref}

\definecolor{acpink}{RGB}{180,72,135}
\definecolor{acpurple}{RGB}{120,80,180}
\definecolor{acgray}{RGB}{119,119,119}

\hypersetup{
  colorlinks=true,
  linkcolor=acpurple,
  urlcolor=acpurple,
  pdfauthor={Jeffrey Alan Scudder},
  pdftitle={LACMA Art + Technology Lab 2026: Aesthetic Computer Budget},
}

\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}

\setlength{\parindent}{0em}
\setlength{\parskip}{0.6em}
\RaggedRight

\newcommand{\ac}{Aesthetic{\color{acpink}.}Computer}

\begin{document}
\thispagestyle{empty}

\vspace*{0.2em}
\begin{center}
  {\acbold\fontsize{26pt}{30pt}\selectfont Aesthetic{\color{acpink}.}Computer}\\[10pt]
  {\large Personal Computers Are Not Done Yet}\\[18pt]
  {\color{acgray}\small Jeffrey Alan Scudder \ $\cdot$ \ LACMA Art + Technology Lab 2026 \ $\cdot$ \ Detailed Project Budget}
\end{center}

\vspace{1em}
{\color{acpink}\hrule height 1pt}
\vspace{1.4em}

\renewcommand{\arraystretch}{1.35}
\begin{tabularx}{\textwidth}{@{}Xr@{}}
  Artist fee (24 months, Fall 2026 to Fall 2028) & \$22,000 \\
  Studio hardware (dev machines, displays) & \$3,500 \\
  AC Device Library fleet (12 AC Blank laptops × \$128 flashed with Aesthetic.Computer, incl.\ cases + shipping + returns) & \$2,000 \\
  Mini Rig portable speakers for installation (5 × \$120) & \$600 \\
  USB drives, cables, peripherals & \$500 \\
  Installation fabrication (furniture, mounts) & \$2,500 \\
  Workshop materials (printed guides, KidLisp reference cards) & \$1,200 \\
  \textbf{2027 Symposium ``boot the cohort''} (cohort demo USB kit, on-site workshop station, travel) & \$2,500 \\
  \textbf{2028 Demo Day ``play the room''} (multi-station install setup, public-program support, v1.0 release) & \$3,000 \\
  \textbf{Paid event helpers} (workshop TAs, install + take-down at Symposium and Demo Day) & \$3,000 \\
  Server + compute infrastructure (hosting, CDN, CI/CD) & \$3,500 \\
  Documentation production (video, photography, translation) & \$2,000 \\
  Contingency (10\%) & \$3,700 \\
  {\color{acgray}\hrulefill} & {\color{acgray}\hrulefill} \\[-0.3em]
  \textbf{Total Requested} & \textbf{\$50,000} \\
\end{tabularx}
\renewcommand{\arraystretch}{1.0}

\vspace{1.6em}

{\color{acgray}\small The 10\% contingency line handles budget slip; paid helpers are a direct expense budgeted for the cohort's two public events. Other funding: open-source sponsorship via GitHub Sponsors and Liberapay supplements ongoing infrastructure costs (hosting, domains, CI). Personal / studio funds cover primary ongoing support.}

\vspace{2em}

{\color{acpink}\hrule height 1pt}
\vspace{1em}

{\bfseries Milestones \& disbursement by phase}

\renewcommand{\arraystretch}{1.3}
\begin{tabularx}{\textwidth}{@{}lXr@{}}
  \textbf{Fall 2026 to Spring 2027} & \textbf{Library Build-Out \& Curriculum.} Establish the AC Device Library (12 AC Blanks, lending-fleet infrastructure, waitlist software), design the 3-level KidLisp workshop curriculum, produce printed reference cards. & \$10,000 \\
  \textbf{Spring to Summer 2027} & \textbf{First Library Cohort + Pilots.} Open the library waitlist; 2 pilot Flash Days at LACMA; first public borrowing cycle; prototype multi-station installation; v0 open-source build guide. & \$8,000 \\
  \textbf{Fall 2027} & \textbf{2027 Symposium ``We boot the cohort.''} Cohort artists and library members flash AC Blanks together and take them home. Public KidLisp workshop; talk alongside the 2023 cohort. & \$7,000 \\
  \textbf{Winter 2027 to Summer 2028} & \textbf{Library Scaling + Extended Workshops.} Kiosk-mode hardening, library curriculum v1 (EN + ES), 4 additional workshops, returns-and-repair cycle for the fleet, complete documentation of the waitlist and lending system. & \$15,000 \\
  \textbf{Fall 2028} & \textbf{2028 Demo Day ``We play the room.''} Multi-station Aesthetic.Computer installation premieres on the LACMA floor; Family Play afternoons open the library fleet to visitors of any age; v1.0 of the open-source build pipeline ships, so any institution can run its own room. & \$10,000 \\
  {\color{acgray}\hrulefill} & {\color{acgray}\hrulefill} & {\color{acgray}\hrulefill} \\[-0.3em]
  & \textbf{Total Requested} & \textbf{\$50,000} \\
\end{tabularx}
\renewcommand{\arraystretch}{1.0}

\vfill
\begin{center}
  {\color{acgray}\small \href{https://aesthetic.computer}{aesthetic.computer} \ $\cdot$ \ \href{mailto:mail@aesthetic.computer}{mail@aesthetic.computer}}
\end{center}

\end{document}
