% !TEX program = xelatex
% Sample meeting in the cards layout — exercises every macro in
% ac-meeting-cards.sty so we can verify the template compiles.

\documentclass[11pt]{article}

\usepackage{fontspec}
\usepackage{unicode-math}
\setmainfont{Latin Modern Roman}[
  Extension=.otf,
  UprightFont=lmroman10-regular,
  BoldFont=lmroman10-bold,
  ItalicFont=lmroman10-italic,
  BoldItalicFont=lmroman10-bolditalic,
]

\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=acpink,
  urlcolor=acpink,
  pdftitle={jeffrey x scott — sample meeting},
}

\usepackage{ac-meeting-cards}

\begin{document}

\meetingcover{jeffrey x scott}{2026-05-30 \enspace$\cdot$\enspace 53 min \enspace$\cdot$\enspace jeffrey, scott}{2026-05-30-1600-jeffrey-x-scott.wav}

\begin{keyideas}
  \keyidea{Decision}{Ship trancepenta stems Friday, unmastered, bare project export.}
  \keyidea{Highlight}{Scott's bridge motif idea is the spine of the next section.}
  \keyidea{Action --- JEFFREY}{Send scott the bare project file by Friday end-of-day.}
\end{keyideas}

\section*{opening{\color{acpink}.}}
\cardsubtitle{small talk and framing.}

\turn{jeffrey}{00:00}{Hey Scott — thanks for jumping on. I wanted to talk through where the trancepenta stems land before Friday.}

\turn{scott}{00:18}{Yeah, perfect timing. I was just listening through last night and I have one structural thing I want to float.}

\turn{jeffrey}{00:34}{Go ahead.}

\mhighlight{Scott's bridge motif idea is the spine of the next section — we should orient everything else around it.}

\turn{scott}{00:41}{The bridge — the part that comes in after the second drop — I think the motif you have there could carry the whole next section. Right now it's a transition. If we let it breathe, it becomes the spine.}

\turn{jeffrey}{01:12}{That's a really good read. Let me sit with that.}

\mnote{Sage made a similar point about letting motifs breathe in a chat last week.}

\section*{decisions{\color{acpink}.}}
\cardsubtitle{what we committed to.}

\turn{scott}{02:30}{So timing-wise — when can you get me the bare stems?}

\turn{jeffrey}{02:38}{I can have them Friday. No mastering, just the bare project export.}

\mdecision{Ship trancepenta stems Friday, unmastered, bare project export.}{50/50 split, publishing his}

\maction{jeffrey}{Send scott the bare project file with no mastering by Friday end-of-day.}

\section*{tangents{\color{acpink}.}}
\cardsubtitle{the small detours.}

\turn{scott}{04:15}{Random question — have you been keeping up with the whistlegraph stuff lately?}

\turn{jeffrey}{04:22}{Yeah, big push this month. I'll show you the recap when we wrap.}

\mquote{Big push this month. I'll show you the recap when we wrap.}

\mfreeform{ask scott about that synth patch from 2023}

\mskipped{1 min 12 sec}

\turn{scott}{05:51}{OK, sending you my notes on the bridge after this.}

\turn{jeffrey}{05:58}{Perfect — talk soon.}

\section*{notes for the future{\color{acpink}.}}
\cardsubtitle{what to revisit next time.}

\turn{jeffrey}{06:00}{For the followup we should plan out a second pass once the stems are in.}

\end{document}
