% !TEX program = xelatex
\documentclass[10pt,letterpaper,twocolumn]{article}

% === GEOMETRY ===
\usepackage[top=0.75in, bottom=0.75in, left=0.75in, right=0.75in]{geometry}

% === FONTS ===
\usepackage{fontspec}
\usepackage{unicode-math}
\setmainfont{Latin Modern Roman}
\setsansfont{Latin Modern Sans}
\setmonofont{Latin Modern Mono}[Scale=0.85]
\newfontfamily\acbold{ywft-processing-bold}[
  Path=../../system/public/type/webfonts/,
  Extension=.ttf
]
\newfontfamily\aclight{ywft-processing-light}[
  Path=../../system/public/type/webfonts/,
  Extension=.ttf
]

% === PACKAGES ===
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{ragged2e}
\usepackage{microtype}
\usepackage{natbib}
\usepackage[colorspec=0.92]{draftwatermark}

% === COLORS ===
\definecolor{acpink}{RGB}{180,72,135}
\definecolor{acpurple}{RGB}{120,80,180}
\definecolor{acdark}{RGB}{64,56,74}
\definecolor{acgray}{RGB}{119,119,119}
\definecolor{draftcolor}{RGB}{180,72,135}

% === DRAFT WATERMARK ===
\DraftwatermarkOptions{
  text=WORKING DRAFT,
  fontsize=3cm,
  color=draftcolor!18,
  angle=45,
  pos={0.5\paperwidth, 0.5\paperheight}
}

% === HYPERREF ===
\hypersetup{
  colorlinks=true,
  linkcolor=acpurple,
  urlcolor=acpurple,
  citecolor=acpurple,
  pdfauthor={@jeffrey},
  pdftitle={notepat.com: From Keyboard Toy to System Front Door},
}

% === SECTION FORMATTING ===
\titleformat{\section}
  {\normalfont\bfseries\normalsize\uppercase}
  {\thesection.}
  {0.5em}
  {}
\titlespacing{\section}{0pt}{1.2em}{0.3em}

\titleformat{\subsection}
  {\normalfont\bfseries\small}
  {\thesubsection}
  {0.5em}
  {}
\titlespacing{\subsection}{0pt}{0.8em}{0.2em}

% === HEADER/FOOTER ===
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[C]{\footnotesize\color{draftcolor}\textit{Working Draft --- not for citation}}
\fancyfoot[C]{\footnotesize\thepage}

% === LIST SETTINGS ===
\setlist[itemize]{nosep, leftmargin=1.2em, itemsep=0.1em}
\setlist[enumerate]{nosep, leftmargin=1.2em}

% === COLUMN SEPARATION ===
\setlength{\columnsep}{1.8em}

% === PARAGRAPH SETTINGS ===
\setlength{\parindent}{1em}
\setlength{\parskip}{0.3em}

% Hyphenation for narrow two-column layout
\tolerance=800
\emergencystretch=1em
\hyphenpenalty=50

\newcommand{\acdot}{{\color{acpink}.}}
\newcommand{\ac}{\textsc{Aesthetic.Computer}}
% Random caps for Aesthetic.Computer branding
\newcount\acrandtmp
\newcommand{\acrandletter}[2]{%
  \acrandtmp=\uniformdeviate 2\relax
  \ifnum\acrandtmp=0\relax#1\else#2\fi%
}
\newcommand{\acrandname}{%
  \acrandletter{a}{A}\acrandletter{e}{E}\acrandletter{s}{S}\acrandletter{t}{T}%
  \acrandletter{h}{H}\acrandletter{e}{E}\acrandletter{t}{T}\acrandletter{i}{I}%
  \acrandletter{c}{C}{\color{acpink}.}\acrandletter{c}{C}\acrandletter{o}{O}%
  \acrandletter{m}{M}\acrandletter{p}{P}\acrandletter{u}{U}\acrandletter{t}{T}%
  \acrandletter{e}{E}\acrandletter{r}{R}%
}

\begin{document}

\twocolumn[{%
\begin{center}
{\acbold\fontsize{24pt}{28pt}\selectfont\color{acdark} notepat{\color{acpurple}.}{\color{acpink}com}}\par
\vspace{0.2em}
{\aclight\fontsize{11pt}{13pt}\selectfont\color{acpink} From Keyboard Toy to System Front Door}\par
\vspace{0.6em}
{\normalsize\href{https://prompt.ac/@jeffrey}{@jeffrey}}\par
{\small\color{acgray} Aesthetic.Computer}\par
{\small\color{acgray} ORCID: \href{https://orcid.org/0009-0007-4460-4913}{0009-0007-4460-4913}}\par
\vspace{0.3em}
{\small\color{acpurple} \url{https://notepat.com} \enspace$\cdot$\enspace \url{https://aesthetic.computer}}\par
\vspace{0.6em}
\rule{\textwidth}{1.5pt}
\vspace{0.5em}
\end{center}

\begin{center}
{\small\color{draftcolor}\textbf{[ working draft --- not for citation ]}}
\end{center}
\vspace{0.3em}

\begin{quote}
\small\noindent\textbf{Abstract.}
\texttt{notepat} began in June 2024 as a 200-line chromatic keyboard sketch inside \ac{}~\citep{scudder2026ac}. Over 401 commits and 20 months it grew to 7{,}903 lines in the browser and 987 lines on bare metal, accumulating five waveform types, room reverb, analog pressure sensitivity, a bare-metal Linux port that boots as PID~1, a branded domain (\texttt{notepat.com}), a dedicated test harness, an alpha Ableton Live device, a VST3 plugin scaffold, and six derivative pieces. I trace that growth: what \texttt{notepat} started as, what features it gained, what infrastructure it pulled into existence, and how a single piece came to function as the strategic front door of a larger creative computing system. This trajectory is a \emph{political economy of a piece}---the feedback loops between a piece's growing capabilities, the platform decisions it motivates, and the resources it attracts over time.
\end{quote}
\vspace{0.5em}
}]

\section{Introduction}

Creative coding platforms are typically evaluated at the system level: language design, editor affordances, community size~\citep{reas2007processing,mccarthy2015p5js,resnick2009scratch}. Even within digital musical instrument research, the unit of analysis is usually the instrument class or interaction paradigm rather than the evolution of a single artifact over years~\citep{magnusson2010designing,marquez2018dmi}. I take a narrower and longer lens: one piece---\texttt{notepat}---inside \ac{}, studied across 20 months of active development.

\texttt{notepat} is a melodic keyboard instrument. McLuhan's observation that the medium shapes the message~\citep{mcluhan1964understanding} applies with particular force: the QWERTY keyboard is not a neutral input device but a medium whose physical layout---rows, clusters, gaps between keys---determines what melodies are ergonomic and what harmonies fall naturally under the hands. The instrument inherits the biases of a typing interface and turns them into musical affordances. It is directly addressable as \texttt{aesthetic.computer/notepat} and through the branded domain \texttt{notepat.com}. As of March 2026, it is the most complex piece in the system, the default boot piece on bare-metal hardware, and the primary subject of performance testing infrastructure. None of this was planned at the outset. The instrument grew, and the platform grew around it.

\section{Origins (June--July 2024)}

\texttt{notepat} has antecedents in two earlier projects by the same author. No Paint (2020--2021), a pixel art tool with six successive repositories, demonstrated that a single-purpose creative piece could build a community of non-technical users. The Whistlegraph performance tools (2019--2023)---real-time subtitle renderers, border notation systems, digital painting delivery services built for live touring with Goodiepal \& Pals---established the practice of building bespoke instruments inside a web runtime. \texttt{notepat} inherits both: No Paint's commitment to a single creative verb, and Whistlegraph's treatment of the browser as a performance instrument.

The first commit appears on June~27, 2024, under the name \texttt{notepad}---a typo that was later reinterpreted as a portmanteau of ``note'' and ``pad.'' The initial version was a minimal keyboard sampler: QWERTY keys mapped to chromatic notes, a single sine waveform, no UI beyond colored rectangles for active keys.

Within five days the piece had gained:
\begin{itemize}
  \item Synth fade-out (25\,ms linear decay to prevent clicks),
  \item Documentation references in the project index,
  \item The corrected name \texttt{notepat}.
\end{itemize}

At this stage the piece was roughly 200 lines. It used the standard \ac{} piece lifecycle (\texttt{boot}, \texttt{paint}, \texttt{act}) and the platform's built-in \texttt{sound.synth} API. The design was deliberately simple: one memorable noun, one immediate playable instrument, no configuration required. This early minimalism is consistent with what \citet{magnusson2010designing} describes as productive constraint---a bounded search space that focuses both player and developer.

\section{Feature Growth}

\subsection{Audio engine (2024--2025)}

The synthesis surface expanded across several axes driven directly by playing needs. This is the dynamic \citet{mcpherson2020idiomatic} describe in computer music languages: tools acquire ``idiomatic patterns'' that reflect and reinforce the aesthetic preferences of their primary users.

\begin{itemize}
  \item \textbf{Waveforms}: sine $\rightarrow$ triangle, sawtooth, square, noise (selectable via keyboard shortcut or parameter).
  \item \textbf{Room reverb}: three-tap delay line (120\,ms intervals), 0.55 feedback coefficient, wet/dry mix controllable via trackpad X-axis slider. Per-sample smoothing on bare metal ($\alpha \approx 5 \times 10^{-5}$ at 192\,kHz) eliminates zipper noise.
  \item \textbf{Pitch shift}: continuous $\pm$1 octave via trackpad Y-axis, with per-sample exponential smoothing ($\alpha \approx 3 \times 10^{-4}$ at 192\,kHz).
  \item \textbf{Octave control}: 9 octaves (1--9), navigable via arrow keys.
  \item \textbf{Quick mode}: fast attack/decay toggle for staccato playing.
  \item \textbf{Metronome}: tempo-synced beat indicator (20--300 BPM), visual flash on downbeats.
\end{itemize}

\subsection{Input and hardware (2024--2026)}

Three hardware integration efforts were driven by \texttt{notepat}'s expressivity requirements. Digital musical instruments are described by \citet{tahiroglu2020probes} as ``probes'' that change how we think about the computational and physical substrate they run on. \texttt{notepat} probed in exactly this way:

\begin{enumerate}
  \item \textbf{NuPhy Air60 HE analog keyboard}: WebHID protocol reverse-engineered from USB captures. Sixty-four-byte activation commands; 16-bit pressure reports per key; velocity-weighted synthesis. Required \texttt{CONFIG\_HIDRAW=y} in the bare-metal kernel.
  \item \textbf{Touchpad as expression controller}: X-axis controls reverb wet mix, Y-axis controls pitch shift. These mappings were designed for \texttt{notepat}'s playing style and have not been generalized.
  \item \textbf{HDMI secondary display}: solid-color output driven by blended note colors during performance. Added to \texttt{ac-native} in March 2026, motivated by live use.
\end{enumerate}

\subsection{Visual design}

The interface evolved from plain colored rectangles to a layered system:

\begin{itemize}
  \item 24-pad split grid (4$\times$3 left octave + 4$\times$3 right octave),
  \item Status bar: clock (LA time with DST), waveform, octave, sample rate, volume, battery, WiFi antenna icon,
  \item Mini piano keyboard visualization with octave shading and black-key dot notation,
  \item HUD label in MatrixChunky8 font with \texttt{.com} superscript branding,
  \item Background tint reactive to time of day (dawn/midday/afternoon/sunset/night),
  \item Note color system: each semitone mapped to a hue (C~=~red through B~=~violet), active notes blend into background.
\end{itemize}

\subsection{DAW integration: Ableton Live}

In January 2025, an alpha Max for Live device (\texttt{AC Notepat.amxd}) was created to embed \texttt{notepat} inside Ableton Live. The device uses Max~9's \texttt{jweb\textasciitilde} object---an audio-enabled embedded web browser---to load the \texttt{notepat} piece and route its Web Audio output directly into Ableton's mixer via \texttt{plugout\textasciitilde}.

This integration required solving several problems not present in the browser or bare-metal contexts:

\begin{itemize}
  \item \textbf{Offline bundling}: a self-contained HTML bundle generator (\texttt{bundle-html.js}) recursively discovers all ES module dependencies, inlines font glyphs, and produces a gzip-compressed self-extracting HTML file with a virtual file system (VFS) and fetch interceptor.
  \item \textbf{AudioWorklet in PACK mode}: the standard AudioWorklet loading path requires a server; the bundled version creates blob URLs from the VFS to load a pre-bundled speaker worklet (\texttt{speaker-bundled.mjs}, $\sim$50\,KB).
  \item \textbf{MIDI-to-keyboard translation}: a planned mapping layer converts MIDI note-on/off events to \texttt{notepat}'s keyboard event model, with CC1 (mod wheel) controlling reverb mix and pitch bend controlling slide mode.
\end{itemize}

A parallel VST3 plugin effort (\texttt{ac-vst/}) scaffolded a C++ Steinberg SDK plugin with a WebView-based UI, targeting a more portable distribution path than Max for Live. Both efforts remain in alpha; the audio engine is not yet functional in the bundled offline context due to worklet loading constraints.

The DAW integration represents a fifth feedback loop in \texttt{notepat}'s political economy: \textbf{distribution pressure $\rightarrow$ bundling infrastructure}. The desire to make \texttt{notepat} available inside professional music production environments forced the platform to develop offline HTML bundling, VFS emulation, and worklet pre-bundling capabilities that no other piece had required.

\subsection{Network and system integration}

\begin{itemize}
  \item WiFi UI: fullscreen network scanner with signal-strength bars, SSID picker, password entry---built into the piece itself for bare-metal contexts where no OS UI exists.
  \item Dark mode: automatic switch at 7\,pm/7\,am LA time, with matching DST calculation in both C and JavaScript.
  \item Projector and recital modes for performance contexts.
\end{itemize}

\section{The Bare-Metal Port}

In late 2024, \texttt{notepat} was ported to \texttt{ac-native}: a custom Linux runtime that boots from USB as PID~1 with no operating system. The bare-metal version (987 lines) runs on QuickJS, renders via DRM/KMS dumb buffers with 3$\times$ nearest-neighbor CPU upscale, and synthesizes audio through direct ALSA \texttt{hw:} access at 192\,kHz.

Embedding a real-time audio synthesizer into a bare-metal Linux environment is an active area of research. \citet{turchet2021elk} describe Elk Audio OS---a Linux-based OS using Xenomai real-time kernel extensions for sub-millisecond audio latency on embedded hardware---as the state of the art in this space. The \texttt{ac-native} approach differs: rather than a general-purpose embedded audio OS, it is a single-piece runtime with no RTOS layer, relying instead on large ALSA period buffers (128 samples at 192\,kHz) and a software-only synthesis engine with 32-voice polyphony.

This port was not incidental. The entire \texttt{ac-native} project---custom kernel configuration, initramfs build pipeline, ALSA audio engine, evdev input handling, WiFi management via \texttt{wpa\_supplicant}, DRM/KMS display---was built primarily to run \texttt{notepat} on physical hardware as a dedicated instrument. The piece justified the platform.

The boot sequence reflects this commitment: after kernel init, \texttt{ac-native} displays a rainbow-animated title, then loads \texttt{notepat.mjs} as the default piece. The instrument is playable within approximately 2 seconds of power-on.

\section{Derivatives and Reuse}

Six artifacts derive directly from \texttt{notepat}:

\begin{enumerate}
  \item \textbf{\texttt{autopat.mjs}} (214 lines): autoplay wrapper with a built-in jukebox (Twinkle, Bach excerpt, scales). Imports and delegates to \texttt{notepat}'s core lifecycle functions.
  \item \textbf{\texttt{notepat-tv.mjs}} (55 lines): UDP-driven remote visualizer for art installations. Receives note data and drives turtle graphics on a projected surface.
  \item \textbf{\texttt{notepat-convert.mjs}} (103 lines): notation translation library; converts single-letter keyboard shortcuts to extended melody syntax. Reused by \texttt{clock}, \texttt{stample}, and other music pieces.
  \item \textbf{Bare-metal \texttt{notepat.mjs}} (987 lines): standalone reimplementation for QuickJS, sharing keyboard mappings and UI conventions but adapted for direct hardware access.
  \item \textbf{\texttt{AC Notepat.amxd}} (Max for Live device): embeds \texttt{notepat} inside Ableton Live via \texttt{jweb\textasciitilde}, routing Web Audio synthesis directly into the DAW mixer. Alpha, January 2025.
  \item \textbf{\texttt{ac-vst} plugin} (C++/Steinberg SDK): VST3 scaffold with WebView UI for DAW-agnostic distribution. Includes MIDI input handling, audio bus routing, and parameter automation stubs.
\end{enumerate}

\section{Infrastructure Pulled into Existence}

\subsection{Routing and branding}

The domain \texttt{notepat.com} is rewritten at the edge to the \texttt{notepat} slug in the primary index function. This required dedicated host-matching logic, custom Open Graph metadata, and a branded HUD element visible during play. \texttt{notepat} is the only piece in the system with its own domain. The \texttt{.com} superscript visible in the instrument's HUD encodes this dual identity into the performance surface itself.

\subsection{Testing infrastructure}

\texttt{notepat} has disproportionate testing investment relative to other pieces:

\begin{itemize}
  \item \textbf{Latency harness} (\texttt{artery/test-notepat-latency.mjs}): measures keyboard-to-sound latency via Chrome DevTools Protocol; reports mean, median, p95, p99. Current measurement: 417\,ms (target: $<$30\,ms).
  \item \textbf{Stability harness} (\texttt{artery/test-notepat-stability.mjs}): 5-minute fuzz test at configurable intensity (2--20 notes/sec); monitors heap growth, sound dictionary size, latency degradation. Latest result: 3{,}229 notes, $-$14\% heap, 0 warnings.
  \item 14 total test scripts across \texttt{artery/}, \texttt{tests/}, and \texttt{.vscode/tests/}.
\end{itemize}

The latency gap (417\,ms vs.\  30\,ms target) represents an open research problem. It is consistent with known browser audio scheduling constraints and the cost of Web Audio API event dispatch from JavaScript. This gap does not arise in the bare-metal version, where ALSA \texttt{hw:} access at 192\,kHz and direct evdev polling yield sub-10\,ms onset latency.

\subsection{Piece-hit analytics}

The piece-hit tracking system (\texttt{/api/piece-hit}) was built in December 2025. It records page loads server-side to MongoDB, distinguishing system and user pieces. In the initial telemetry window (2025-12-31 through 2026-01-02), \texttt{notepat} recorded 27 hits and ranked 4th among 44 tracked built-ins. Current instrumentation limitations---no typed namespaces, no authenticated attribution, fire-and-forget POST calls that were being killed by the Netlify Functions runtime before completion---mean that the full picture of \texttt{notepat} usage since 2024 is not captured. The fire-and-forget bug was identified and patched in March 2026.

\section{Political Economy of a Piece}

Kurzweil's ``law of accelerating returns'' describes how technological systems accumulate capability at an accelerating rate as each advance enables the next~\citep{kurzweil2005singularity}. At the scale of a single creative piece, a similar dynamic is visible. \texttt{notepat}'s trajectory illustrates feedback loops between a piece and its host platform that \citet{nieborg2018platformization} describe at the level of cultural industries: platform affordances shape production, and production demands reshape platforms.

\begin{enumerate}
  \item \textbf{Feature pressure $\rightarrow$ platform capability}: \texttt{notepat} needed room reverb, so the audio engine gained delay lines. It needed analog pressure, so \texttt{ac-native} gained HID raw support. It needed a secondary display, so HDMI output was added to the DRM backend.
  \item \textbf{Branding pressure $\rightarrow$ routing decisions}: the desire for a memorable URL led to \texttt{notepat.com}, which required edge-function host rewriting and custom social previews.
  \item \textbf{Quality pressure $\rightarrow$ testing investment}: a latency regression in \texttt{notepat} motivated the construction of a general-purpose piece-testing harness (artery), which now benefits all pieces.
  \item \textbf{Hardware pressure $\rightarrow$ entire runtime}: the ambition to run \texttt{notepat} on dedicated hardware justified building \texttt{ac-native}: custom kernel, initramfs, ALSA engine, DRM display, WiFi stack---thousands of lines of C.
  \item \textbf{Distribution pressure $\rightarrow$ bundling infrastructure}: the desire to embed \texttt{notepat} inside Ableton Live forced the platform to develop offline HTML bundling with a virtual file system, fetch interceptor, and pre-bundled AudioWorklets---infrastructure no other piece had required.
\end{enumerate}

\citet{born2015music} observe that music technology investment is never neutral---it reflects and reinforces the aesthetic priorities, social positions, and resources of those who build and fund it. The \texttt{notepat} trajectory makes this visible at the scale of an individual developer and a single open-source instrument: sustained personal investment in one piece materially shapes what is technically possible for every other piece in the system.

Live coding communities have developed an explicit vocabulary for this kind of relationship between practice and tool~\citep{blackwell2022livecoding}: the instrument and the instrumentalist co-evolve. \texttt{notepat}'s 401 commits are one record of that co-evolution in a non-live-coding context---a persistent instrument that accumulated capability through use rather than through upfront design.

\section{Current Scale}

Table~\ref{tab:scale} summarizes \texttt{notepat}'s footprint as of March 8, 2026.

\begin{table}[h]
\small
\centering
\begin{tabularx}{\columnwidth}{l r}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Commits (browser version) & 401 \\
Browser LOC & 7{,}903 \\
Bare-metal LOC & 987 \\
Derivative pieces & 6 \\
Total ecosystem LOC & 9{,}262 \\
Waveform types & 5 \\
Dedicated test scripts & 14 \\
Files mentioning ``notepat'' in repo & 159 \\
Age (first commit) & June 27, 2024 \\
\bottomrule
\end{tabularx}
\caption{notepat scale indicators (March 8, 2026).}
\label{tab:scale}
\end{table}

\section{Conclusion}

\texttt{notepat} demonstrates that a single piece can function as the strategic center of gravity for a creative computing platform. Its growth from a 200-line keyboard sketch to a 9{,}000-line ecosystem with its own domain, hardware runtime, DAW integration, and test infrastructure was not planned---it emerged from sustained use and the feedback loops described above.

The methodological contribution is a unit of analysis: the \emph{piece trajectory}. Studying a single piece across its full commit history---what features it gained, what infrastructure it pulled into existence, what derivatives it spawned---reveals platform dynamics invisible at the system level. Digital musical instrument research has examined DMI longevity as a problem of adoption~\citep{marquez2018dmi} and DMIs as probes of computational substrate~\citep{tahiroglu2020probes}. The piece trajectory combines both perspectives with a political economy lens: instruments accumulate capability, and that accumulation has material consequences for the systems that host them.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
