% !TEX program = xelatex
% media-estate.tex — overview deck: the AC × whistlegraph media estate.
% First deck in the papers "decks" lane: 16:9 landscape, one page per slide.
\documentclass[12pt]{article}

% === GEOMETRY === (16:9, screen-first)
\usepackage[paperwidth=338.667mm, paperheight=190.5mm,
            left=20mm, right=20mm, top=14mm, bottom=12mm]{geometry}

% === FONTS ===
\usepackage{fontspec}

\setmainfont{Latin Modern Roman}[
  Extension=.otf,
  UprightFont=lmroman10-regular,
  BoldFont=lmroman10-bold,
  ItalicFont=lmroman10-italic,
  BoldItalicFont=lmroman10-bolditalic,
]
\setmonofont{Latin Modern Mono}[
  Extension=.otf,
  UprightFont=lmmono10-regular,
  ItalicFont=lmmono10-italic,
  Scale=0.85,
]

% Display face — title text only (punctuation glyphs unreliable).
\newfontfamily\acbold{ywft-processing-bold}[
  Path=../../system/public/type/webfonts/,
  Extension=.ttf
]

% === COLORS ===
\usepackage{xcolor}
\definecolor{acpink}{RGB}{180,72,135}
\definecolor{acpurple}{RGB}{120,80,180}
\definecolor{acdark}{RGB}{40,36,48}
\definecolor{acgray}{RGB}{119,119,119}
\definecolor{acred}{RGB}{200,50,75}
\definecolor{acgreen}{RGB}{40,140,70}
\definecolor{acamber}{RGB}{200,120,20}
\definecolor{acbg}{RGB}{250,250,250}

% === PACKAGES ===
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{ragged2e}
\usepackage{microtype}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  urlcolor=acpurple,
  pdfauthor={@jeffrey},
  pdftitle={The Media Estate},
}

\pagestyle{empty}
\pagecolor{acbg}
\color{acdark}
\setlength{\parindent}{0pt}
\renewcommand{\arraystretch}{1.45}

% Status words for the access matrices.
\newcommand{\okfull}{{\color{acgreen}\textbf{full}}}
\newcommand{\okpart}{{\color{acamber}\textbf{partial}}}
\newcommand{\oknone}{{\color{acred}\textbf{none}}}
\newcommand{\rb}{{\color{acred}\textbullet}\hspace{0.6em}}

\newcommand{\ac}{Aesthetic\textcolor{acpink}{.}Computer}

% One slide = one page. #1 section label, #2 title (YWFT-safe words only).
\newenvironment{slide}[2]{%
  \clearpage
  {\footnotesize\color{acgray}\MakeUppercase{#1}\hfill}%
  {\footnotesize\color{acgray}\ac{} \textperiodcentered{} @jeffrey}\par
  \vspace{0.1em}
  {\acbold\fontsize{31pt}{35pt}\selectfont\color{acdark}#2\par}
  \vspace{0.9em}
  \large
}{%
  \vfill
  {\hfill\footnotesize\color{acgray}\thepage}%
}

\begin{document}

% ── 1 · title ────────────────────────────────────────────────────────
\vspace*{\fill}
\begin{center}
  {\acbold\fontsize{58pt}{62pt}\selectfont\color{acdark}the media estate\par}
  \vspace{1.2em}
  {\fontsize{19pt}{24pt}\selectfont\itshape\color{acpink}
    one funnel, nine surfaces --- whistlegraph \textrightarrow{} \ac\par}
  \vspace{2.2em}
  {\large\color{acgray} @jeffrey \textperiodcentered{} July 2026 \textperiodcentered{} planning hub: \texttt{/social} in the monorepo\par}
\end{center}
\vspace*{\fill}
{\hfill\footnotesize\color{acgray}\thepage}

% ── 2 · the shape ────────────────────────────────────────────────────
\begin{slide}{the shape}{two brands, one funnel}
\textbf{@whistlegraph} is the established audience; \textbf{@aesthetic.computer} is the growth target.
The funnel runs one way --- ``made with aesthetic.computer'' posts, never an audience merge.

\vspace{1.2em}
\begin{center}
\begin{tabular}{@{}lll@{}}
\toprule
 & \textbf{whistlegraph} & \textbf{aesthetic.computer} \\
\midrule
TikTok & 2.6M followers \textperiodcentered{} 99M likes & 391 followers \\
YouTube & 5{,}340 subs \textperiodcentered{} 2.1M views & 4 subs \textperiodcentered{} 70 views \\
Instagram & (established; count TODO) & under 1k --- \textit{the gate that matters} \\
Posting since & dormant since \textasciitilde 2023 & just beginning \\
\bottomrule
\end{tabular}
\end{center}

\vspace{1.2em}
\rb The audience already exists --- the whistlegraph archive still moves \textbf{\textasciitilde 48{,}000 video views per day} with zero new posts.\\
\rb The problem is not reach. It is that \textbf{the pipe between the brands is not connected}, and most surfaces cannot be posted to by script yet.
\end{slide}

% ── 3 · the accounts ─────────────────────────────────────────────────
\begin{slide}{inventory}{twelve accounts, nine surfaces}
Tracked in \texttt{social/accounts.json}; strategy and tooling index in \texttt{social/SCORE.md}.

\vspace{0.8em}
\begin{center}
\normalsize
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Account} & \textbf{Surface} & \textbf{Role} & \textbf{Standing} \\
\midrule
@aesthetic.computer & Instagram & \textbf{primary growth target} & under 1k \\
@aesthetic.computer & TikTok & primary growth target & 391 followers \\
Aesthetic Dot Computer & YouTube & music + readings archive & 4 subs \\
@aesthetic.computer & Bluesky & changelog + paintings feed & 81 followers \\
AC Readings & Podcast & essays \textrightarrow{} Spotify/Apple/YT & stats TODO \\
@promptDOTac & X & dormant brand outlet & revive by mirror \\
Aesthetic Dot Computer & Spotify artist & distribution via DistroKid & claim blocked \\
@whistlegraph & TikTok & \textbf{funnel source} & 2.6M followers \\
@whistlegraph & YouTube & funnel source & 5.3k subs \\
@whistlegraph & Instagram & funnel source & count TODO \\
@whistlegraph & X & dormant funnel & count TODO \\
(jeffrey) & Are.na & curation surface & token not vaulted \\
\bottomrule
\end{tabular}
\end{center}
\end{slide}

% ── 4 · what the archive says ────────────────────────────────────────
\begin{slide}{evidence}{what the archive says}
Top whistlegraph YouTube videos (of 210 uploads, 2.1M total views, pulled 2026-07-05):

\vspace{0.8em}
\begin{center}
\begin{tabular}{@{}rll@{}}
\toprule
\textbf{Views} & \textbf{Video} & \textbf{Era} \\
\midrule
245k & mommy wow WHISTLING GRAPHIX & 2021 lo-fi chalk originals \\
213k & zigzag x circle, i love mommy wow & 2021 lo-fi chalk originals \\
108k & The Longest Whistlegraph Ever & 2022 relatable-caption era \\
83k  & seasons changing & 2022 relatable-caption era \\
65k  & hey, are you from LA? & 2022 relatable-caption era \\
\bottomrule
\end{tabular}
\end{center}

\vspace{1em}
\rb The \textbf{lo-fi, process-visible originals} massively outperform --- the audience came for watching a form being made, not for polish.\\
\rb Nothing posted since 2023, and the numbers held. \textbf{Dormant is not dead.}\\
\rb This is the exact premise \ac{} content should lead with: \textit{the making, not the made.}
\end{slide}

% ── 5 · access today ─────────────────────────────────────────────────
\begin{slide}{audit}{access today}
Can we post to it by script, from the monorepo, right now?

\vspace{0.8em}
\begin{center}
\normalsize
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Surface} & \textbf{Access} & \textbf{Via} & \textbf{Gap} \\
\midrule
YouTube (both channels) & \okfull & \texttt{toolchain/youtube/yt.mjs} & quota: 6 uploads/day \\
Bluesky & \okfull & \texttt{at/} scripts, app password & --- \\
Podcast (Buzzsprout) & \okfull & \texttt{reading} skill & stats pull \\
Papers / essays & \okfull & self-hosted, no gatekeeper & --- \\
Instagram (both) & \okpart & read-only unofficial sessions & \textbf{no publish at all} \\
TikTok (both) & \okpart & OAuth app in silo; ingest only & unaudited: no publish \\
Spotify (music) & \okpart & DistroKid semi-automated & artist claim blocked \\
Are.na & \okpart & proven gig scripts & token not vaulted \\
X (both) & \oknone & --- & no developer app \\
\bottomrule
\end{tabular}
\end{center}

\vspace{0.9em}
The scripted surfaces are the quiet ones; \textbf{the loud surfaces are all manual}. That asymmetry is the whole build-out.
\end{slide}

% ── 6 · the keystone ─────────────────────────────────────────────────
\begin{slide}{the keystone}{instagram pays three ways}
One build-out --- official publishing via the Instagram API with Instagram Login
(no Facebook Page; own accounts publish in dev mode, no app review) --- unlocks three
separate outcomes:

\vspace{1.1em}
\begin{enumerate}[leftmargin=2.2em, itemsep=0.7em]
  \item \textbf{The 1k gate.} Scripted reels at cadence are the only realistic path to
        1{,}000 followers on @aesthetic.computer --- which unlocks \textbf{Trial Reels}
        (test on non-followers, promote winners). A hard gate, not a soft one.
  \item \textbf{The funnel.} Same tooling posts ``made with aesthetic.computer''
        re-cuts to @whistlegraph --- the 2.6M-audience bridge, finally scriptable.
  \item \textbf{The Spotify claim.} The Spotify for Artists verification has been
        \textbf{blocked since May} on one thing: a trancenwaltz post live on the
        @aesthetic.computer grid. First API-published post un-blocks the waiting
        draft reply (\texttt{pop/spotify-for-artists-claim-reply.md}).
\end{enumerate}

\vspace{1.1em}
Detail that shapes the tooling: IG publish takes a \textbf{public video URL}, not a file
upload --- our DO Spaces CDN feeds it, so the CLI pushes to Spaces first.
\end{slide}

% ── 7 · where content lands ──────────────────────────────────────────
\begin{slide}{supply}{the mills already run}
Every surface has a content mill behind it in the monorepo --- production is not the bottleneck.

\vspace{0.8em}
\begin{center}
\normalsize
\begin{tabular}{@{}lll@{}}
\toprule
\textbf{Pipeline} & \textbf{Produces} & \textbf{Lands on} \\
\midrule
\texttt{pop/} music mill & tracks + visualizers + 9:16 cuts & DistroKid, YouTube, reels \\
\texttt{marketing/kidlisp-reels/} & KidLisp process reels, 9:16 & IG + TikTok at-bats \\
\texttt{recap/} & narrated vertical video & IG + TikTok \\
\texttt{papers/} \textrightarrow{} readings & essays + spoken audio & site, Buzzsprout, Spotify, Apple \\
\texttt{at/} scripts & changelog threads, paintings & Bluesky (X mirror next) \\
whistlegraph archive & 950+ TikToks, 193 videos & re-cut fodder for funnel posts \\
\bottomrule
\end{tabular}
\end{center}

\vspace{0.9em}
\rb One edit, many surfaces: the same 9:16 short goes to IG Reels and TikTok every time.\\
\rb The archive is a \textbf{library, not a museum} --- top performers become funnel re-cuts.
\end{slide}

% ── 8 · the plan ─────────────────────────────────────────────────────
\begin{slide}{build-out}{api access, in value order}
\begin{enumerate}[leftmargin=2.2em, itemsep=0.55em]
  \item \textbf{Instagram official publishing} --- Meta app + Instagram Login; then
        \texttt{ig.mjs} in the toolchain, same shape as \texttt{yt.mjs}. \textit{The keystone.}
  \item \textbf{X developer app} --- free tier covers a mirror cadence;
        \texttt{x.mjs} posts the same short drops as Bluesky.
  \item \textbf{TikTok Content Posting audit} --- the OAuth app already exists in silo;
        an approved audit makes the 2.6M-follower account scriptable.
  \item \textbf{Are.na token to the vault} --- scripts already speak the API.
  \item \textbf{Buzzsprout stats} --- wire downloads into \texttt{accounts.json} snapshots.
  \item \textbf{YouTube quota bump} --- when cadence outgrows six uploads a day.
\end{enumerate}

\vspace{1em}
Done already: YouTube both channels (2026-07-05) \textperiodcentered{} Bluesky \textperiodcentered{}
Buzzsprout publish \textperiodcentered{} papers.

\vspace{0.6em}
Every unlock lands as a CLI in the same idiom --- vault-held tokens,
\texttt{--as <channel>} identity switching, private-first defaults.
\end{slide}

% ── 9 · the strategy ─────────────────────────────────────────────────
\begin{slide}{doctrine}{lead with the making}
\textbf{Positioning.} \ac{} is the hand-built-instrument brand: \textit{``I made the tool
that made this.''} Visible authorship --- KidLisp coming together, a synth answering live ---
inoculates against the AI-slop read. The archive evidence (slide 4) says this is also
simply what the audience rewards.

\vspace{0.9em}
\textbf{Cadence.} 4--7 process-over-product shorts a week; below 1k a flop costs nothing ---
every reel is an at-bat. One repeatable series people can anticipate.

\vspace{0.9em}
\textbf{Funnel discipline.} @whistlegraph stays its own voice; it tags, it never merges.
Music ships as Aesthetic Dot Computer; papers and readings ship as @jeffrey.

\vspace{0.9em}
\textbf{Never:} follow/unfollow churn, engagement pods, bought followers --- they poison
the non-follower targeting that Trial Reels and the algorithm both depend on.
\end{slide}

% ── 10 · next ────────────────────────────────────────────────────────
\begin{slide}{next}{four portal sessions}
Each unlock needs one short human session at a developer portal --- everything after is scripted:

\vspace{0.9em}
\begin{enumerate}[leftmargin=2.2em, itemsep=0.55em]
  \item \textbf{Meta app} (developers.facebook.com) --- connect both IG professional
        accounts, hand back app id, secret, and two long-lived tokens. \textit{\textasciitilde 15 min.}
  \item \textbf{X developer account} (developer.x.com) --- create the app under the
        brand owner. \textit{\textasciitilde 10 min.}
  \item \textbf{TikTok audit application} --- submit the Content Posting API review
        on the existing silo app. \textit{\textasciitilde 15 min + wait.}
  \item \textbf{Are.na token} --- generate once, vault it. \textit{\textasciitilde 2 min.}
\end{enumerate}

\vspace{1.1em}
Planning hub: \texttt{social/SCORE.md} + \texttt{social/accounts.json} ---
append a dated snapshot whenever fresh numbers are pulled; every new tool gets a row
in the index so \texttt{/social} stays the single front door.

\vspace{0.9em}
{\color{acgray}\small This deck: \texttt{papers/deck-media-estate/} --- the first entry
in the papers decks lane.}
\end{slide}

\end{document}
