% !TEX program = xelatex
\documentclass[manuscript,anonymous,review]{acmart}

\settopmatter{printacmref=false}
\renewcommand\footnotetextcopyrightpermission[1]{}
\pagestyle{plain}

\begin{document}

\title{Technical Demonstration Requirements}
\subtitle{Aesthetic Computer --- ACM C\&C 2026 Demo}

\author{Anonymous}
\affiliation{\institution{Anonymous Institution}}

\maketitle

\section{Space and Setup}

\begin{itemize}
  \item \textbf{Table space:} Standard demo table (approximately 180cm $\times$ 75cm).
  \item \textbf{Display:} One external monitor (24" or larger), provided by the authors if not available. HDMI input preferred.
  \item \textbf{Laptop:} Authors will bring a laptop running Chrome to drive the main display.
  \item \textbf{Audio:} Small powered speaker or headphones for the demo station. Musical pieces (notepat, tone) produce audio that is part of the demonstration.
  \item \textbf{Power:} Two power outlets (laptop + monitor or speaker).
\end{itemize}

\section{Network Requirements}

\begin{itemize}
  \item \textbf{Internet connection:} Required. The system loads from a public URL. Wi-Fi is sufficient.
  \item \textbf{Bandwidth:} Minimal after initial load ($<$1 MB cached). Real-time features (chat, multiplayer) use WebSocket connections with low bandwidth overhead.
  \item \textbf{Ports:} Standard HTTPS (443) and WSS (443). No special firewall rules needed.
  \item \textbf{Fallback:} If internet is unreliable, the demo can run from a local development server on the presenter's laptop. Attendees would connect via a local Wi-Fi hotspot.
\end{itemize}

\section{Attendee Participation}

\begin{itemize}
  \item Attendees use \textbf{their own phones, tablets, or laptops}---no dedicated hardware required.
  \item The system works in any modern browser (Chrome, Safari, Firefox, Edge). No app installation or account creation is needed.
  \item A printed reference card will list starting commands and a QR code for the URL.
  \item Participation is drop-in: attendees can join or leave at any time during the demo session.
\end{itemize}

\section{Accessibility}

\begin{itemize}
  \item The system accepts keyboard, touch, mouse, and gamepad input.
  \item Text rendering uses high-contrast colors on solid backgrounds.
  \item The prompt interface is compatible with screen readers (standard HTML input element).
  \item Audio is not required for the visual/drawing portions of the demo.
\end{itemize}

\section{Setup and Teardown}

\begin{itemize}
  \item \textbf{Setup time:} 10 minutes (open laptop, connect display, verify internet).
  \item \textbf{Teardown time:} 5 minutes.
  \item \textbf{No special installation:} The entire system runs in a web browser. No software needs to be installed on venue machines.
\end{itemize}

\section{Presenter Attendance}

One author will be present throughout the demo session to guide attendees, explain the system, and facilitate collaborative moments (multi-user drawing and music sessions).

\end{document}
